Parsing Postfix log files

### 2009-01-13 doc/logparser-other-parsers-review.tex
Further corrections after proofreading.

@@

Berkeley DB only supports storing simple \textbf{(key, value)} pairs,
unlike \gls{SQL} databases that store arbitrary tuples.  In \gls{LMA}'s
main table the key is an integer referred to by secondary tables, and the
value is a \gls{CSV} line containing all of the data for that row.  The
secondary by-sender, by-recipient, by-date, and by-\gls{IP} tables use the
sender/recipient/date/\gls{IP} as the key, and the value is a \gls{CSV}
list of integers referring to the main table.  This effectively
re-implements \gls{SQL} foreign keys, but without the functionality offered
by even the most basic of \gls{SQL} databases (joins, ordering, searches,
etc.).  It also requires custom code to search on some combination of the
above, though the authors of \gls{LMA} did provide some queries: IP-STORY,
FROM-STORY, DAILY-EMAIL, and DAILY-REJECT\@.  Berkeley DB appears to be the
least useful of the three output formats: it does not provide the
functionality of a basic \gls{SQL} database, and unlike \gls{CSV} files it
can not be used with standard Unix text processing tools.

@@

    \item \url{http://www.thesawmill.co.uk/formats/postfix.html} \newline{}
        Fields extracted: from, to, server, UID, relay, status, number of
        recipients, origin hostname, origin \gls{IP}, and virus.  It also
        counts the number of and total size of all mails delivered.  The
        fields \texttt{server}, \texttt{uid}, \texttt{relay}, and
        \texttt{virus} are not explained in the documentation:
        \texttt{server} is probably the hostname or \gls{IP} of the server
        the mail is delivered to; \texttt{relay} might be the delivery
        method: \gls{SMTP}, local delivery, or \gls{LMTP}; \texttt{uid}
        might be the uid of the user submitting mail locally.  Postfix does
        not perform any form of virus checking (though it has many options
        for cooperating with an external virus scanner), so the
        \texttt{virus} field is a mystery.

@@

When parsing Postfix log files Splunk parses the standard
syslog\glsadd{syslog} fields at the beginning of the log line, and extracts
any \texttt{key=value} pairs occurring after the standard syslog prologue:
to and from addresses, HELO hostname, and protocol (\gls{SMTP} or
\gls{ESMTP}).  \parsername{} extracts noticeably more data (client and
server \gls{IP} and hostname, queueid, start and end times, timestamps for
each log line, \gls{SMTP} code, and message ID), though it does not make
the full text of the line available (this could be trivially added if
desired, but would greatly increase the size of the resulting database).
The full power of \gls{SQL} is available when searching the data extracted
by \parsername{}, allowing the user to search on arbitrarily complicated
conditions.

@@

@@

AWStats will produce simple graphs for many different services, but
supporting many different services without special purpose code limits its
functionality.  The data it will extract from an \gls{MTA} log file is
limited in comparison to \parsername{}: time2, email, email\_r, host,
host\_r, method, url, code, and bytesd.  There is no explanation for any of
those fields in the documentation (\parsername{} provides copious
documentation), so the author could not understand the extracted data, nor
determine what data is missing in comparison to \parsername{}.  AWStats
coerces Postfix log files into Apache\footnote{The Apache web server is the
most popular HTTP server in use over the past 10 years; more information is
available at \url{http://httpd.apache.org/}.} format log files, for
analysis by AWStats' HTTP log file parser.  The converting parser only
deals with a small portion of the log lines generated by Postfix, silently
skipping those it cannot deal with, and does not distinguish between
different types of rejection; it would be a lot of work to extend it to
handle new log lines.  Although it does correlate log lines by queueid (not
by pid), it does not deal with any of the other complications described in
\sectionref{complications}.  AWStats supports saving data but the format of
the saved data is not documented, as far as the author could tell.  It also
supports reading compressed log files, but that functionality was not
tested.

When tested with the \numberOFlogFILES{} test log files AWStats' reported
that it parsed 9,240,075 (88.70\%) of 10,416,129 log lines, skipping
1,176,050 (11.29\%) corrupt log lines; however there are
\numberOFlogLINES{} lines in the \numberOFlogFILES{} log files, so AWStats
parsed only 17.15\% of the input lines, ignoring the remaining 82.85\%.

@@

AWStats was not a suitable base for this project, because it assumes that
all log files can be rewritten to be compatible with web server log files,
and will contain similar data; coercing Postfix log files into web server
log files, without substantial data loss, would require fully parsing the
Postfix log files, so AWStats would not be required.  It may be possible to
use AWStats graphing capabilities to generate reports, by generating log
files to use as input to AWStats from the data extracted by \parsername{}.

@@

Anteater does not have any English documentation so it is impossible for
this author to accurately comment on the analysis it performs.  It did not
run successfully when tested, and its parsing would certainly be out of
date as Postfix has evolved considerably since this tool was last updated
(2003/11/06).  As it neither ran successfully nor has documentation the
author can read a detailed review cannot be provided.

@@

YAALA uses a plugin-based system to analyse log files and produce HTML
output reports, with all the parsing and report generation handled by
modules.  Using YAALA as a base would be only slightly less work than
starting from scratch, as both the input and output modules would need to
be written specially; it may even be more work to implement the parser
within the constraints of YAALA\@.  YAALA supports storing previously
gathered data using Perl's Storable module~\cite{perl-storable}, so other
Perl programs can use Storable to load, examine, and optionally modify the
data; \parsername{} uses a well documented database which is accessible
from the majority of programming languages.  This information was gleaned
from the source code, as the documentation is sadly lacking.

YAALA provides a Postfix parser that extracts the following two types of
fields from specific log lines:

@@

\noindent{}YAALA's Postfix parser extracts some of the fields \parsername{}
does: it stores either the \gls{IP} or the hostname for client and server,
not both; it omits the HELO hostname, queueid, \gls{SMTP} code, size of
each accepted mail, start and end times, timestamps for each log line, and
message ID\@.  It extracts some data that \parsername{} does not: the delay
in delivering each mail, and how many times delivery was deferred for each
mail; these could easily be extracted by \parsername{} if desired.  XXX
EXTRACT DELAY AND DEFER ONCE AUTOMATIC EXTRACTION HAS BEEN FINISHED\@.
Unlike \parsername{}, YAALA does not maintain separate counters for each
restriction; this rules out the possibility of using the collected data for
optimisation, testing or understanding of restrictions.  YAALA's Postfix
parser does not deal with the complications explained in
\sectionref{complications}, though it does correlate log lines by queueid.

@@

\noindent{}Examining the source code reveals that the parser looks for
\texttt{<key>=<value>} pairs in each log line, extracts them, and
correlates the data by queueid.  This approach will find the following
fields: HELO hostname, queueid, \gls{SMTP} code, sender and recipient
addresses, and size of accepted mails.  It is unclear if the parser will
extract any further data.  Lire misses the following fields extracted by
\parsername{}: client and server \gls{IP} and hostname, start and end
times, timestamps of each log line, and message ID\@.

Lire supports multiple output formats for generated reports (text, HTML,
PDF, and Excel 95) but the reports do not appear to be customisable;
\parsername{} does not produce any reports.  Lire's report contains less
detail than Pflogsumm, and is considerable harder to configure.  Lire
supports saving extracted data for later report generation, but accessing
this data from another application is undocumented; given the source code
it should be possible, with enough time and effort, to understand the
format.  \parsername{} uses an \gls{SQL} database to make accessing the
extracted data as easy as possible.

@@

It also counts the number of log lines parsed and skipped.  It omits client
\gls{IP} and hostname, server \gls{IP}, HELO hostname, queueid, and message
ID\@.  It extracts the delay for delivered mails, which \parsername{} does
not.  Log lines are correlated based on the queueid (referred to as
sessionname [sic] within Logrep), but not by \gls{pid}.  The parsing is
error prone: empty fields are saved when the log line does not match the
regex, though it appears that they will not overwrite existing data.  Most
notably rejections are completely ignored, making it unsuitable for the
purposes of this project.  It does not try to address any of the
complications in \sectionref{complications} except for correlating by
queueid.

Logrep does not come with any documentation, though some scant
documentation is available on its website (\parsername{} provides copious
documentation).  It requires a web browser to interact with it, so
automated log file processing will be difficult, whereas enabling automated
processing is a key part of \parsernames{} design.  Sadly all the author's
attempts to use Logrep failed, as it was unable to access the log files
selected; this appears to be a bug rather than operator error.  If it was
caused by operator error, the interface needs improvement as the (minimal)
instructions were followed as closely as possible, and multiple attempts
were made.  As parsing failed it was not possible to review the reports
Logrep can generate (available in HTML only), or to examine the
(undocumented) format in which it can save extracted data for subsequent
reuse.

@@

There are other programs available which perform basic Postfix log file
parsing (some to a greater level of detail than others), but few attempt to
correlate log lines by queueid (none correlate by \gls{pid}) to produce an
overall record of the journey of each mail through Postfix.  None of the
reviewed parsers collect the breadth of information gathered by
\parsername{}, or make it as easy to extend the parser to handle new log
lines.  Most of the parsers generate a report and immediately discard the
data extracted from the log files; those that do not discard the data
typically retain it in a format inaccessible to other tools.  Nearly all of
the parsers reviewed can produce a report of greater or lesser detail and
complexity, unlike \parsername{}.  The quality of the documentation offered
by the subset of parsers that provide some varies from unusable to good;
none of the parsers provide any documentation on the format of their data
stores (if they have one).  Fewer than half of the parsers were capable of
parsing the \numberOFlogFILES{} test log files, and improving or extending
parsing would have been quite a difficult task for any of the parsers.
Table \refwithpage{Summary of parsers' features} provides a summary of the
parsers' features.

The overriding difference between \parsername{} and the other parsers
reviewed herein is that none of them aim for the high level of
understanding of Postfix log files achieved by \parsername{}.

\begin{table}[htb]
    \caption{Summary of parsers' features}
    \empty{}\label{Summary of parsers' features}
    \begin{tabular}{llllll}
        \tabletopline{}%
        Parser          & Parsed test   & Data              & Custom            & Documentation  & Source       \\
                        & log files?    & store?            & reports?          & quality?       & code?        \\
        \tablemiddleline{}%
        \gls{LMA}       & No            & Yes               & No                & Poor           & Yes          \\ 
        Pflogsumm       & Yes           & No                & Partial \dag{}    & Good           & Yes          \\
        Sawmill         & Yes           & Yes               & Searches          & Very good      & \nialpha{}   \\
        Splunk          & Yes           & Yes               & Searches          & Abundant       & No           \\
                        &               &                   & \& reports        & but poor       &              \\
        Isoqlog         & No            & Yes               & No                & Poor           & Yes          \\
        AWStats         & Partially     & Yes               & Partial \dag{}    & Good           & Yes          \\
        Anteater        & No            & No                & No                & Poor           & Yes          \\
        YAALA           & No            & Yes \ddag{}       & Searches          & Poor           & Yes          \\
        Lire            & Yes           & Yes               & Yes               & Reasonable     & Yes          \\
        Logrep          & No            & Yes               & No                & Poor           & Yes          \\
        \parsername{}   & Yes           & Yes \nibeta{}     & No \nichi{}       & XXX            & Yes          \\
        \tablebottomline{}%
    \end{tabular}

>>>>

XXX DO ANY OF THE PARSERS DISTINGUISH BETWEEN DIFFERENT TYPES OF
REJECTIONS\@?

@@

Berkeley DB only supports storing simple \textbf{(key, value)} pairs,
unlike \gls{SQL} databases that store arbitrary tuples.  In \gls{LMA}'s
main table the key is an integer referred to by secondary tables, and the
value is a \gls{CSV} line containing all of the data for that row.  The
secondary by-sender, by-recipient, by-date, and by-\gls{IP} tables use the
sender/recipient/date/\gls{IP} address as the key, and the value is a
\gls{CSV} list of integers referring to the main table.  This effectively
re-implements \gls{SQL} foreign keys, but without the functionality offered
by even the most basic of \gls{SQL} databases (joins, ordering, searches,
etc.).  It also requires custom code to search on some combination of the
above, though the authors of \gls{LMA} did provide some queries: IP-STORY,
FROM-STORY, DAILY-EMAIL, and DAILY-REJECT\@.  Berkeley DB appears to be the
least useful of the three output formats: it does not provide the
functionality of a basic \gls{SQL} database, and unlike \gls{CSV} files it
can not be used with standard Unix text processing tools.

@@

    \item \url{http://www.thesawmill.co.uk/formats/postfix.html} \newline{}
        Fields extracted: from, to, server, UID, relay, status, number of
        recipients, origin hostname, origin \gls{IP}, and virus.  It also
        counts the number of and total size of all mails delivered.  The
        fields \texttt{server}, \texttt{uid}, \texttt{relay}, and
        \texttt{virus} are not explained in the documentation:
        \texttt{server} is probably the hostname or \gls{IP} address of the
        server the mail is delivered to; \texttt{relay} might be the
        delivery method: \gls{SMTP}, local delivery, or \gls{LMTP};
        \texttt{uid} might be the uid of the user submitting mail locally.
        Postfix does not perform any form of virus checking (though it has
        many options for cooperating with an external virus scanner), so
        the \texttt{virus} field is a mystery.

@@

When parsing Postfix log files Splunk parses the standard
syslog\glsadd{syslog} fields at the beginning of the log line, and extracts
any \texttt{key=value} pairs occurring after the standard syslog prologue:
to and from addresses, HELO hostname, and protocol (\gls{SMTP} or
\gls{ESMTP}).  \parsername{} extracts noticeably more data (client and
server \gls{IP} address and hostname, queueid, start and end times,
timestamps for each log line, \gls{SMTP} code, and message ID), though it
does not make the full text of the line available (this could be trivially
added if desired, but would greatly increase the size of the resulting
database).  The full power of \gls{SQL} is available when searching the
data extracted by \parsername{}, allowing the user to search on arbitrarily
complicated conditions.

@@

XXX DOCUMENTATION\@?

@@

AWStats can produce simple graphs from many different services' log files,
but supporting numerous log files formats without special purpose code
limits its functionality.  The data it can extract from an \gls{MTA} log
file is limited in comparison to \parsername{}: time2, email, email\_r,
host, host\_r, method, url, code, and bytesd.  There is no explanation for
any of those fields in the documentation (in contrast \parsername{}
documents everything it extracts), so the author could not understand what
data is extracted, nor determine what data is missing in comparison to
\parsername{}.  AWStats coerces Postfix log files into Apache\footnote{The
Apache web server is the most popular HTTP server in use over the past 10
years; more information is available at \url{http://httpd.apache.org/}.}
format log files, for analysis by AWStats' HTTP log file parser.  The
converting parser only deals with a small portion of the log lines
generated by Postfix, silently skipping those it cannot deal with, and does
not distinguish between different types of rejection; extending it to parse
all log lines would be at least as much work as writing a new parser.  It
does correlate log lines by queueid (not by pid), but it does not deal with
any of the other complications described in \sectionref{complications}.
AWStats supports saving data, but the format of the data store is not
documented.  It also supports reading compressed log files, but that
functionality was not tested.

When tested with the \numberOFlogFILES{} test log files AWStats' reported
that it parsed 9,240,075 (88.70\%) of 10,416,129 log lines, skipping
1,176,050 (11.29\%) corrupt log lines; however there are
\numberOFlogLINES{} lines in the \numberOFlogFILES{} log files, so AWStats
parsed only 15.21\% of the input lines, declared 1.93\% were corrupt, and
ignored the remaining 82.85\%.

@@

AWStats was not a suitable base for this project, because it assumes that
all log files can be rewritten to be compatible with web server log files,
and will contain similar data; coercing Postfix log files into web server
log files, without substantial data loss, would require fully parsing the
Postfix log files without using AWStats, i.e.\ would require writing a
separate parser.  It may be possible to use AWStats' graphing capabilities
to generate reports, by generating input for AWStats from the data
extracted by \parsername{}.

@@

Anteater does not have any English documentation so it is impossible for
this author to accurately comment on the analysis it performs.  It did not
run successfully when tested, and its parsing would certainly be out of
date because Postfix has evolved considerably since this tool was last
updated (2003/11/06).  As it neither ran successfully nor has documentation
the author can read, a detailed review cannot be provided.

@@

YAALA uses a plugin-based system to analyse log files and produce reports
in HTML format, with all the parsing and report generation handled by
plugins.  Using YAALA as a base for this project would be only slightly
less work than starting from scratch, as both the input and output modules
would need to be written specially; it may be more work to implement a
parser within the constraints of YAALA\@.  YAALA supports storing
previously gathered data using Perl's Storable module, so other Perl
programs can use Storable to load, examine, and optionally modify the data;
\parsername{} uses a well documented database which is accessible from the
majority of programming languages.  This information was gleaned from the
source code, as the documentation is sadly lacking.

YAALA provides a Postfix parser that extracts the following types of fields
from specific log lines:

@@

\noindent{}YAALA's Postfix parser extracts some of the fields \parsername{}
does: for client and server it stores either the \gls{IP} address or the
hostname, not both; it omits the HELO hostname, queueid, \gls{SMTP} code,
size of each accepted mail, start and end times, timestamps for each log
line, and message ID\@.  It extracts some data that \parsername{} does not:
the delay in delivering each mail, and how many times delivery was deferred
for each mail; these could easily be extracted by \parsername{} if desired.
XXX DELAY AND DEFERRAL\@.  Unlike \parsername{}, YAALA does not maintain
separate counters for each restriction; this rules out the possibility of
using the collected data for optimisation, testing, or understanding of
restrictions.  YAALA's Postfix parser does not deal with the complications
explained in \sectionref{complications}, though it does correlate log lines
by queueid.

@@

\noindent{}Examining the source code reveals that the parser looks for
\texttt{<key>=<value>} pairs in each log line, extracts them, and
correlates the data by queueid.  This approach will extract the following
data: HELO hostname, queueid, \gls{SMTP} code, sender and recipient
addresses, and size of accepted mails.  It is unclear if the parser will
extract any further data.  Lire misses the following fields extracted by
\parsername{}: client and server \gls{IP} address and hostname, start and
end times, timestamps of each log line, and message ID\@.

Lire supports multiple output formats for generated reports (text, HTML,
PDF, and Excel 95) but the reports do not appear to be customisable;
\parsername{} does not produce any reports.  Lire's report contains less
detail than Pflogsumm, and it is considerable harder to configure.  Lire
supports saving extracted data for later report generation, but the format
of this data store is undocumented; given the source code it should be
possible, with enough time and effort, to understand the format.
\parsername{} uses an \gls{SQL} database to make accessing the extracted
data as easy as possible.

@@

It also counts the number of log lines parsed and skipped.  It omits client
\gls{IP} address and hostname, server \gls{IP}, HELO hostname, queueid,
timestamps of each log line, and message ID\@.  It extracts the delay for
delivered mails, which \parsername{} does not.  Log lines are correlated
based on the queueid (referred to as sessionname [sic] within Logrep), but
not by \gls{pid}.  The parsing is error prone: empty fields are saved when
the log line does not match the regex, though it appears that they will not
overwrite existing data.  Most notably rejections are completely ignored,
making it unsuitable for the purposes of this project.  It does not try to
address any of the complications in \sectionref{complications} except for
correlating by queueid.

Logrep does not come with any documentation, though some scant
documentation is available on its website (\parsername{} provides copious
documentation).  It requires a web browser to interact with it, so
automated log file processing will be difficult, whereas enabling automated
processing is a key part of \parsernames{} design.  Sadly all the author's
attempts to use Logrep failed, as it was unable to access the log files
selected; this appears to be a bug rather than operator error.  If it was
caused by operator error, the interface needs improvement as the (minimal)
instructions were followed as closely as possible, and multiple attempts
were made.  Because parsing failed it was not possible to review the
reports Logrep can generate (available in HTML only), or to examine the
(undocumented) format in which it can save extracted data for subsequent
reuse.

@@

There are other programs available which perform basic Postfix log file
parsing (some to a greater level of detail than others).   None of the
reviewed parsers collect the breadth of information gathered by
\parsername{}, or make it as easy to extend the parser to handle new log
lines.  Some correlate log lines by queueid (none correlate by pid); none
deal with any of the other complications described in XXX REFERENCE
complications.  All of the reviewed parsers generate a report, and some
provide a greater or lesser degree of customisation.  Most have a data
store, but only \gls{LMA} provides any documentation on its format; some
deliberately make the data store inaccessible to other tools.  Most but not
all of the parsers provide documentation, with the quality varying from
unusable to very good.  Fewer than half of the parsers were capable of
parsing the \numberOFlogFILES{} test log files, and improving or extending
parsing would have been quite a difficult task for any of the parsers.
Table \refwithpage{Summary of parsers' features} provides a summary of the
parsers' features.

XXX REPHRASE The overriding difference between \parsername{} and the other
parsers reviewed herein is that none of them aim for the high level of
understanding of Postfix log files achieved by \parsername{}.

\begin{table}[htb]
    \caption{Summary of parsers' features}
    \empty{}\label{Summary of parsers' features}
    \begin{tabular}{llllll}
        \tabletopline{}%
        Parser          & Parsed test   & Data              & Custom            & Documentation  & Source       \\
                        & log files?    & store?            & reports?          & quality?       & code?        \\
        \tablemiddleline{}%
        \gls{LMA}       & No            & Yes               & No                & Poor           & Yes          \\ 
        Pflogsumm       & Yes           & No                & Partial \dag{}    & Good           & Yes          \\
        Sawmill         & Yes           & Yes               & Searches          & Very good      & \nialpha{}   \\
        Splunk          & Yes           & Yes               & Searches          & Abundant       & No           \\
                        &               &                   & \& reports        & but poor       &              \\
        Isoqlog         & No            & Yes               & No                & No English     & Yes          \\
                        &               &                   &                   & documentation  &              \\
        AWStats         & Partially     & Yes               & Partial \dag{}    & Good           & Yes          \\
        Anteater        & No            & No                & No                & Poor           & Yes          \\
        YAALA           & No            & Yes \ddag{}       & Searches          & Poor           & Yes          \\
        Lire            & Yes           & Yes               & Yes               & Reasonable     & Yes          \\
        Logrep          & No            & Yes               & No                & None           & Yes          \\
        \parsername{}   & Yes           & Yes \nibeta{}     & No \nichi{}       & XXX            & Yes          \\
        \tablebottomline{}%
    \end{tabular}
